---
title: "Chap6"
output: html_document
---
```{r }
library (dplyr)
library (tidyverse)
library (ggplot2)
library (tidyr)
library (corrplot)
library (lme4)
```
Data sets were prepared in meet_and_repeat.R file

**Chapter 8 in MABS: analysis of longitudinal data : Graphical displays and summary measure approach using **RATS** data

```{r }
#read data
RATSL<- read.csv ("rats.csv", row.names = 1)
# raw graphic

fig1 <- ggplot(RATSL, aes(x = Time, y = Weight, group = ID))+ geom_text(aes(label = Group))+ scale_x_continuous(name = "Time (days)", breaks = seq(0, 60, 10)) + scale_y_continuous(name = "Weight (grams)") + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
fig1
fig2 <- ggplot(RATSL, aes(x = Time, y = Weight, group = ID)) +
  geom_line()
fig2
```

*** Data description
Data was extracted from a nutrition study conducted in three groups of rats (Crowder and Hand, 1990). 

*ID*: number of rats in study

*Group*: three groups in study (1,2,3)

Variables (*weight*): body weight (grams)

Time of study (*WD*): seven Times

**Interpret**
 we want to see the diffecences between the variable *weight* of rats changing during the time in 3 groups.

It can be seen from the fig1 above: group 1 seperate compared group2 and 3. Group 2 and 3 are duplicated, somehow we can guess there is no difference between group 2 and 3.

**Scatterplot matrix of repeated measures in rat growth data with raw data to see independent of variables 
```{r}
RATS <- read.table("https://raw.githubusercontent.com/KimmoVehkalahti/MABS/master/Examples/data/rats.txt", header = TRUE, sep = '\t')
pairs(RATS[, 3:13], cex = 0.7)
```
 a scatterplot matrix of the repeated measures of weight does demonstrate that the repeated measures are certainly not independent of one another

*** Standardise the scores for RAT data set and plot after standardised
```{r}
RATSL <- RATSL %>%
  group_by(Time) %>%
  mutate( stdWeight = (Weight - mean(Weight))/sd(Weight) ) %>%
  ungroup()
glimpse(RATSL)
# Plot again with the standardised bprs
fig1 <- ggplot(RATSL, aes(x = Time, y = Weight, group = ID))+ geom_text(aes(label = Group))+ scale_x_continuous(name = "Time (days)", breaks = seq(0, 60, 10)) + scale_y_continuous(name = "Weight (grams)") + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
fig1
fig2 <- ggplot(RATSL, aes(x = Time, y = Weight, group = ID)) +
  geom_line()
fig2
```

```{r}
# Number of time baseline (Time =1 ) 
n <- RATSL$week %>% unique() %>% length()
# Make a summary data:
RATSLS <- RATSL %>%
  group_by(Group, Time) %>%
  summarise( mean=mean(Weight), se=sd(Weight)/sqrt(n) ) %>%
  ungroup()
```
***See how data set change 

```{r}
glimpse(RATSLS)
p1 <- ggplot(RATSLS, aes(x = factor(Time), y = mean, linetype = Group, shape = Group))
p2 <- p1 + geom_line() + scale_linetype_manual(values = c(1,2,3))
p3 <- p2 + geom_point(size=3) + scale_shape_manual(values = c(1,2,3))
p4 <- p3 + geom_errorbar(aes(ymin=mean-se, ymax=mean+se, linetype="1"), width=0.3)
p5 <- p4 + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
p6 <- p5 + theme(legend.position = c(0.8,0.8))
p7 <- p6 + scale_y_continuous(name = " mean of weights")
p7
```


```{r}
# Make a summary data of based on group during the time
RATSL8S <- RATSL %>%
  filter(Time > 1) %>%
  group_by(Group, ID) %>%
  summarise( mean=mean(Weight) ) %>%
  ungroup()
glimpse(RATSL8S)
p1 <- ggplot(RATSL8S, aes(x = Group, y = mean))
p2 <- p1 + geom_boxplot()
p3 <- p2 + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
p4 <- p3 + stat_summary(fun.y = "mean", geom = "point", shape=23, size=4, fill = "white")
p5 <- p4 + scale_y_continuous(name = "mean(Weight), Time")
p5
```


From above graph, in each group has outliers. So we need to eliminate outliers

```{r}
# Remove the outlier:
RATSL8S1 <- RATSL8S %>%
  filter(mean >= 250  & mean <= 550)
glimpse(RATSL8S1)
p1 <- ggplot(RATSL8S1, aes(x = Group, y = mean))
p2 <- p1 + geom_boxplot()
p3 <- p2 + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
p4 <- p3 + stat_summary(fun.y = "mean", geom = "point", shape=23, size=4, fill = "white")
p5 <- p4 + scale_y_continuous(name = "mean(Weight), Time ")
p5
```

** T-test 
```{r}
# Without the outlier, apply Student's t-test, two-sided:
result <- aov(mean ~ Group, data=RATSL8S1 )
summary(result)
```

**Chapter 9 linear mixed effects models for normal response: using **BPRS** data

```{r }
BPRSL<- read.csv ("bprs.csv", row.names = 1)
summary (BPRSL)
pairs(BPRSL, cex=0.7)
ggplot(BPRSL, aes(x = Time, y = bprs, group = subject)) +
  geom_line()
```
*** Data description:
Data containts:

- Subjects (*subject*): 40 male were randomly assigned to one of two treatment groups 

- Treatment (*treatment*): variable 1 or 2 

- Brief psychiatric rating scale (*BPRS*): score of each subject, which was assessed by the level of 18 symptom constructs such as hostility, suspiciousness, hallucinations and grandiosity...

- week (*week*) from 1 - 8 weeks


1) Have a look on data variables
```{r }
BPRSL_reg <- lm(bprs ~ week + treatment, data = BPRSL)
summary(BPRSL_reg)
```
Look at assocition between Psychiatric rating scale and week treatment and method of treatment, in which coefficients of *week* = - 2.27, and *treatment* = 0.57. The BPRS score seems to be decreasing in both treatment groups,  we are interested whether one treatment seems to be more efficient than the other one.

Look at P-value of treatment  = 0.66, it seems methods of treatment does not effec to model

**** The Random Intercept Model
```{r }
library (lme4)
#Fit the random intercept model with the subject as the random effect
BPRSL_ref <- lmer(bprs ~ week + treatment + (1 | subject), data = BPRSL, REML = FALSE)
summary(BPRSL_ref)
```

****Slippery slopes: Random Intercept and Random Slope Model
```{r }
# create a random intercept and random slope model
BPRSL_ref1 <- lmer(bprs ~ week + treatment + (week | subject), data = BPRSL, REML = FALSE)
summary(BPRSL_ref1)
```


Time to interact: Random Intercept and Random Slope Model with interaction
```{r }
# perform an ANOVA test on the two models
anova(BPRSL_ref, BPRSL_ref1)
```
