---
title: "chapter5"
output: html_document
---

```{r}
library (ggplot2)
library (tidyr)
library (stringr)
library (dplyr)
library (corrplot)
library(GGally)
library(stringr)
library (psych)
```
## Question 1: Graphical overview and summary variables in the data
```{r}
#read data
human <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/human2.txt", sep = ",", header = T)
#look at 10 first values
head (human)
names (human)
str (human)
summary (human)
human$GNI <- str_replace(human$GNI, pattern=",", replace ="") %>% as.numeric()
str(human$GNI)
#plot to see correlation between variables in data
fig1 <- pairs(human, pch = 21,bg = c ("red", "green3", "blue", "purple", "violet", "yellow", "orange","grey"))
fig2 <-cor(human) %>% corrplot.mixed()
```
* variables explanation:
The dataset has 155 observation and 8 variables
*Edu2.FM*: education ratio

*Labo.Fm*: participation ratio

*Edu.Exp*: mean of yeas of education

*Life.Exp*: measures life expectancy at birth

*GNI*: Gross national income per capital

*Mat.Mor*: maternity mortality

*Ado.Birth*: Adolescent of birth rate

*Parli.F*: Percetange of female representatives in parliament

* From the fig1 and fig2, we can see the correlations between the variables.

+ There seems to be strong positive correlations between: *Edu.Exp* and *Life.Exp*, *Mat.Mor* and *Ado.Birth*...

+ A negative correlation: between *Mat.Mor* and *Life.Exp*...


## Question 2: Perform principal component analysis (PCA) on the not standardized human data

```{r}
pca_human.notstand <- prcomp(human)
summary1<- summary (pca_human.notstand)
print (summary1)
# rounded percetanges of variance captured by each PC
pca_pr <- round(100*summary1$importance[2, ], digits = 1)
# create object pc_lab to be used as axis labels
pc_lab <- paste0(names(pca_pr), " (", pca_pr, "%)")
biplot(pca_human.notstand, cex = c(0.8, 1), col = c("grey40", "deeppink2"), xlab = pc_lab[1], ylab = pc_lab[2])
```
From the unstandardized PCA, we can see that the PCs cannot be separated. 

## Question 3: Perform principal component analysis (PCA) on the standardized human data 

```{r}
human_std <- scale(human)
pca_human_std <- prcomp(human_std)
summary2<- summary(pca_human_std)
print (summary2)
pca_pr <- round(100*summary2$importance[2, ], digits = 1)
pca_pr
pc_lab <- paste0(names(pca_pr), " (", pca_pr, "%)")
biplot(pca_human_std, cex = c(0.8, 1), col = c("orange", "green4"), xlab = pc_lab[1], ylab = pc_lab[2])
```




## Question 4: Give personal interpretations of the first two principal component dimensions based on the biplot drawn after PCA on the standardized human data.

In the standardized model the first components (PC1) explain 53.6 % of the variance 
and the second dimension (PC2) explaints 16.2 % of the variance. 
PC3 has 9.5 % of the variance and so on 

## Question 5: Tea dataset analysis
install packages  Factominer first

install.packages("FactoMineR")
```{r}
library(FactoMineR)
library (ggplot2)
library (tidyr)
library (dplyr)
library (corrplot)
library(GGally)
library(stringr)
library (psych)

#calling data tea
data(tea)
#structure of data tea
str(tea)
#dimentions of data
dim (tea)
# tea includes 300 observation with 36 variables
# visualize 10 first variables
fig5 <- pairs(tea[,1:10])
# fig5 shows that correlation between first 10 variables are not linear.. 
# column names to keep in the dataset
keep_columns <- c("Tea", "How", "how", "sugar", "where", "lunch")
# select the 'keep_columns' to create a new dataset
tea_time <- dplyr::select(tea, one_of(keep_columns))
summary(tea_time)
str(tea_time)
head (tea_time)
gather(tea_time) %>% ggplot(aes(value)) + facet_wrap("key", scales = "free") + geom_bar() + theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))
```



### Multiple Correspondence Analysis on the tea data

```{r}
mca <- MCA(tea_time, graph = FALSE)
summary(mca)
par(mfrow=c(1,2))
plot(mca, invisible=c("ind"), title  = "before")
plot(mca, invisible=c("ind"), habillage = "quali", title  = "after")

```
The plot above shows a global pattern within the data. Rows (individuals) are represented by blue points and columns (variable categories) by red triangles.

The distance between any row points or column points gives a measure of their similarity (or dissimilarity). Row points with similar profile are closed on the factor map. The same holds true for column points.
